Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-11-09 analysis/paper/paper.qmd
edit discussion

Our models premise that the choice of weapon-tip types reflects the socio-environmental circumstances. To examine the possible connections between the roles of stemmed points and population dynamics or environmental change during the Late Paleolithic period, we compared the temporal and spatial patterns of TCSA to results of simulated Mean Annual Temperature (MAT) and Summed probability distribution (SPD) from @park2022did. We found clear correlation between stemmed points and temperature (@spa-gis-tcsa-sites and @fig-tcsa-radio-carbon-box, and Figure 9 from @park2022did)
around 4 to 6 ℃

 We consider that TCSA, or the function of the stemmed points, are strongly dependent on the surrounding environment or their situation. @no-figure-no-more: B shows that the temperature slowly decreased until around 22 ka and then went up again with a little range of fluctuations. Temperatures differ significantly between sites, with a maximum difference of 5℃. Results of SPD show several peaks, as indicated by the positive deviations from the three null models (shaded in red), and downs (shaded in blue), but no drastic growth or decline (@no-figure-no-more: C). From the climate and population study, we could observe that there is no dominating change but minor and often fluctuations throughout the Late Paleolithic and variations per site. Overall this pattern is well matched with the TCSA results, indicating the people produced stemmed points as multifunctional tools suitable for unexpected or various situations in their habitats similar to the case of Clovis points (@eren2022plains).

\<--! In discussion we then compare those temporal patterns with demographic and environmental proxies based on summed probability distributions (SPD) and annual temperature changes from @park2022did. --\>

>>>>

Our models premise that the choice of weapon-tip types reflects the socio-environmental circumstances. To examine the possible connections between the roles of stemmed points and population dynamics or environmental change during the Late Paleolithic period, we compared the temporal and spatial patterns of TCSA to results of simulated Mean Annual Temperature (MAT) and Summed probability distribution (SPD) from @park2022did.

Stemmed points were most commonly used in environments with a temperature between 4 and 6 ℃ (@spa-gis-tcsa-sites and @fig-tcsa-radio-carbon-box, and Figure 9 from @park2022did).

\<--!upto here --\>
 We consider that TCSA, or the function of the stemmed points, are strongly dependent on the surrounding environment or their situation. @no-figure-no-more: B shows that the temperature slowly decreased until around 22 ka and then went up again with a little range of fluctuations. Temperatures differ significantly between sites, with a maximum difference of 5℃. Results of SPD show several peaks, as indicated by the positive deviations from the three null models (shaded in red), and downs (shaded in blue), but no drastic growth or decline (@no-figure-no-more: C). From the climate and population study, we could observe that there is no dominating change but minor and often fluctuations throughout the Late Paleolithic and variations per site. Overall this pattern is well matched with the TCSA results, indicating the people produced stemmed points as multifunctional tools suitable for unexpected or various situations in their habitats similar to the case of Clovis points (@eren2022plains).
